### 2023-02-16 tesis_Luis/Chapter3/chapter3.tex
Add files via upload

The precision measurement of the luminosity delivered to the CMS by the LHC is important for a variety of reasons. Online, the luminosity measurement provides realtime feedback on the LHC performance and operation. Offline, the luminosity measurement is a crucial component of physics analysis, either for measuring the cross section of observed processes or for setting upper limits in searches for processes BSM \cite{pas_18}.\\
A total of seven systems are used for measuring luminosity at CMS. These systems are referred to as luminometers.  Each luminometer reads out a rate of the specific quantities observed in the detector (hits, tracks, clusters, etc.). This rate, R, should be proportional to the instantaneous luminosity,$\mathcal{L}_{inst}$, with the constant of proportionality given by the visible cross section $\sigma_{vis}$ \cite{pas_18}:

@@

The determination of $\sigma_{vis}$ is carried out through van der Meer (vdM) scans performed with a dedicated LHC machine setup, exploited by S. Van der Meer for luminosity measurements at ISR \cite{vdM_cal}.

\section{Pixel Cluster Counting method}
The Pixel Cluster Counting method exploits the very large number of pixels in the tracker of the CMS detector. At design luminosity of $10^{34}cm^{-2}s^{-1}$, the detector occupancy is less than 0.1\% on average (the average number of clusters per minimum bias interaction is about 50 and the average number of pixels per cluster is about 5) \cite{pas-13}. The very low occupancy leads to an excellent linear detector response with increasing LHC luminosity \cite{pas_18}. \\
The PCC method uses the rate of pixel clusters in the CMS pixel detector to provide an offline luminosity measurement. Pixel clusters are formed from adjacent pixels, as mentioned in section \ref{pixel_clust_reco}. PCC measurement uses the data collected with the standard CMS trigger system with triggers requiring colliding bunches but not any specific event activity; this is referred as "zero-bias trigger'' \cite{pas_18}.\\
Each bunch crossing gives rise to some number of pp interactions, with each such interaction resulting in some number of pixel clusters. If the average over several zero-bias events is taken, the mean number of pixel clusters per event is \cite{PCC_PAS_12_001}:

@@

\section{Luminosity calibration: van der Meer method}
In the practice, as mentioned in chapter 1, the properties of the colliding beams are not know precisely, such as the bunch density profile, so that the integral in \ref{lumi_2} cannot be solved analytically. In LHC a experimental techinque know as vdM method is implemented  with a dedicated machine setup to bring the bunch density profiles close to Guassian distributions.\\
The VdM scan method, %, exploited by S. Van der Meer for luminosity measurements at ISR  \cite{vdM_cal} and
applied by LHC Experiments in Run 1 and Run 2, allows to measure and evaluate the integral over the proton bunch densities in eq. \ref{lumi_2}. For the vdM scan method to be applicable it is also assumed that the two bunch $\rho$ can be factorized into independent terms in x and y \cite{pas-15}.  The value of the two beam overlap integrals in Eq. (2) is determined by varying the beam separation and measuring the resulting rates \cite{pas_18}:

@@

Experimentally, two separate scans in the x and y directions are performed. The beams are scanned against each other  while measuring the rate (normalized by the product of the beam currents) at a certain number of separation steps, fitting the resulting points with a functional form, and using the fitted function to extract $\Sigma_{x,y}$ and $R(0,0)$ in \ref{sigmavis_eq}. Fig. \ref{vdm_sketch} shows an sketch of the beam positions during vdM scans in X and Y planes together with the detector rate as a function of beam separation.

@@

%In practice, the integral in Eq. (4) is evaluated by performing two separate scans in the x and y directions, measuring the rate (normalized by the product of the beam currents) at a certain number of separation steps, fitting the resulting points with a functional form, and using the fitted function to determine the overall integral [PAS-18]. ADD HEDWING FIGURE. "this is illsutrated in fig N". Once Σx and Σy are determined, then Eq. (1) can be used to obtain the overall visible cross section σvis [pas-18].

%NOTA: Hay un pas, creo,en donde dicen lode beam current normalizado, haciendo referencia a la gráfica. Pero tal vez esto vendría en la siguiente seeción, la de resultados, ya que muestre las fitted graphs.

@@

%DUDA: En la qe. lumi_2 decimos n_b colliding bunches y aquí hablaremos de solo un coliding buch. Lumi inst. se refiere a un solo coliding bunch? Cuál es la diferencia? En el vdM Scan usamos un solo colliding bunch?

\subsection*{Backgrounds}
Before performing the fit to extract $\Sigma_{x}$ and $\Sigma_{y}$, the raw measured rates must be corrected for background in the detector. The background contribution is estimated independently, using a special period known as Super Separation period where the beams are separated, ensuring that the only contribution is from background either beam-induced background or detector noise. There are three sources of beam-induced background: Beam Gas Elastic (BGE), Beam Gas Inelastic (BGI) and Beam Halo (BH). BGE contribution are all coherent and quasi-elastic nuclear elastic and coulomb scattering for multi-turn beam-gas interactions around the ring; typically the interaction of the primary beam proton and the aperture takes place at the collimator. BGI  are all inelastic interactions of primary beam protons with rest gas in the beam pipe. The interaction rate is dominated by the vacuum quality in the various beam line elements upstream CMS; therefore the origin of this contribution is distributed all along the long straight section. BH component is caused by the inefficiency of the main collimation system. Protons can escape from one collimator and being intercepted by another collimator \cite{bkg_source}.\\
%https://cds.cern.ch/record/1316945/files/CR2010_169.pdf
%https://cds.cern.ch/record/2816679/files/document.pdf
% Electronic noise?:
%  https://cds.cern.ch/record/1293521/files/Thesis-2002-Gu.pdf
Several other corrections are applied to the raw data before the vdM fit is performed.\\
%These corrections take into account drifts in the position of the LHC orbit, beam-beam interaction effects, and the beam separation length scale. The intensities are measured by the LHC beam current monitors and are corrected for bunched spurious (so-called “ghost” and “satellite”) charges \cite{pas_18}.

%----------- MOVE TO NEXT CHAPTER-------------%

%Before performing the fit to extract $\Sigma_{x}$ and $\Sigma_{y}$, the raw measured rates must be corrected for background in the detector. The background contribution is estimated independently, using two special periods, each 5 minutes long, during the VdM scan where the beams were separated by 6$\sigma$ (where $\sigma_{b}$ is the beam size) in both $x$ and $y$ directions, ensuring that the only contribution is from background (either beam-induced background or detector noise). The estimated background is then subtracted from the raw data before the fit is performed using a Gaussian-kind  function, which was found in 2018 to adequately describe the measured rates as a function of beam separation.\\
%Several other corrections are applied to the raw data before the VdM fit is performed. These corrections take into account drifts in the position of the LHC%orbit, beam-beam interaction effects, and the beam separation length scale. The event rates are normalized by the bunch intensities before the fit. The intensities are measured by the LHC beam current monitors and are corrected for bunched spurious (so-called “ghost” and “satellite”) charges.

%\section{Data set: 2018 vdm scan program}
%
%For 2018, the VdM scans were performed during LHC fill 6868 on June 30 and July 1 at $\sqrt{s}=13 \text{TeV}$. The LHC filling scheme used  124 colliding bunch pairs at the CMS interaction point (IP5) widely spread over the orbit to reduce long-range beam-beam effects and detector afterglow.  The resulting beam size $\sigma_{b}$ at the beginning of the fill was in the range of approximately 85–95 and 80–90 $\mu$m in x and y, respectively, increasing over time in the x dimension and decreasing over time in the y dimension. No crossing angle was used for collisions. The resulting peak pileup was approximately $\mu$ = 0.6, much lower than in a regular physics fill \cite{pas_18}.\\
%The bunch intensities were approximately 7–9$\times 10^{10}$ protons per filled bunch, resulting in a total beam intensity of slightly above $10^{10}$ protons per beam. The total beam intensities were measured with the DC Current Transformers (DCCT), and the bunch currents were measured with the Fast Beam Current Transformers (FBCT) \cite{pas_18}.\\
%To ensure a dataset with a high event count for PCC even at large beam separations, CMS gated the zero-bias triggers on 5 bunch pairs (BCIDs 265, 865, 1780, 2192, and 3380) and recorded events with a total rate of 27.7 kHz \cite{pas_18}.\\
%The CMS VdM scan program was conducted in two parts (``takes'') due to an alarm. Apart from the vdM scans other studies were carried out. Betwen that studies, only the imaging scans are used to compute $\sigma_{vis}$. The standar vdM scans are referred as ``norm'' and the imaging scans as ``img''. In the first take of the VdM scan program a vdM scan ``norm1'' and a imaging scan ``img1'' were carried out. The second part consisted of the scans ``img2'', ``img3'', ``norm2'', ``norm3'' and ``norm4'' were done. In total are 7 scans to determine $\sigma_{vis}$ \cite{pas_18}.\\
%In the vdM scans the two beams were separated by $6 \sigma_{b} \approx 600 \mu$m and scanned across one another in a sequence of 25 steps with 30 seconds per step. For the beam imaging scans  one beam (beam 1 in the first pair, followed by beam 2 in the second pair) is kept fixed at its nominal position while the other is separated and scanned in 19 steps from  $+$4.5$\sigma_{b}$ to $-$4.5$\sigma_{b}$ with 46 seconds per step. In each of both scans (scan  pairs), the scan was performed first in the x direction and then in the y direction \cite{pas_18}.\\
%Fig. \ref{scan_prog}  shows the beam positions for the two beams in the x and y directions as measured by the DOROS BPMs during the scan program, showing all scan pairs of the 2018 scan program.
%
%\begin{center}
%  \begin{figure}[ht]
%    \centering
%    \includegraphics[scale=.45]{Chapter4/2018Scanprogram.png}
%    \caption[2018 scan program]{Relative change in beam positions measured by the DOROS BPMs during the 2018 scan program for the two individual beams in the horizontal x and vertical y planes, as a function of the elapsed time from the beginning of the program. The top row shows the take 1 of the program before the alarm at CMS, while the bottom two rows show the program after the alarm \cite{pas_18}.}
%    \label{scan_prog}
%  \end{figure}
%\end{center}
%
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

@@

\section{Pixel Cluster Counting method}
 \\

@@

\section{Luminosity calibration: van der Meer method}

@@

@@

@@

\subsection*{Backgrounds}

### 2023-02-16 tesis_Luis/Chapter4/chapter4.tex
Add files via upload

%These corrections take into account drifts in the position of the LHC orbit, beam-beam interaction effects, and the beam separation length scale. The intensities are measured by the LHC beam current monitors and are corrected for bunched spurious (so-called “ghost” and “satellite”) charges \cite{pas_18}.

@@

For 2018, the vdM scans were performed during LHC fill 6868 on June 30 and July 1 at pp collision energy $\sqrt{s}$ of 13 TeV. The LHC filling scheme used  124 colliding bunch pairs at the CMS interaction point (IP5) widely spread over the orbit to reduce long-range beam-beam effects and detector afterglow.  The resulting beam size $\sigma_{b}$ at the beginning of the fill was in the range of approximately 85–95 $\mu$m and 80–90 $\mu$m in x and y, respectively, increasing over time in the x dimension and decreasing over time in the y dimension. No crossing angle was used for collisions. The resulting peak pileup was approximately $\mu$ = 0.6, much lower than in a regular physics fill ($<\mu>$= 32 in 2018 ) \cite{pas_18}.\\ %pileup phys: https://twiki.cern.ch/twiki/bin/view/CMSPublic/LumiPublicResults
The bunch intensities were approximately 7–9$\times 10^{10}$ protons per filled bunch, resulting in a total beam intensity of slightly above $10^{13}$ protons per beam. The total beam intensities were measured with the DC Current Transformers (DCCT), and the bunch currents were measured with the Fast Beam Current Transformers (FBCT) \cite{pas_18}.\\
To ensure a dataset with a high event count for PCC even at large beam separations, CMS gated the zero-bias triggers on 5 bunch pairs: BCIDs (Bunch Crossing ID) 265, 865, 1780, 2192, and 3380; and recorded events with a total rate of 27.7 kHz \cite{pas_18}.\\
The CMS vdM scan program was conducted in two parts due to an alarm. Apart from the vdM scans other studies were carried out called Imaging Scans, which can also be used to measure $\sigma_{vis}$. Imaging scans are performed by moving only one beam. The standard vdM scans are referred as ``norm'' and the imaging scans as ``img''. In the first part of the scan program a vdM scan ``norm1'' and a imaging scan ``img1'' were carried out. The second part consisted of the scans ``img2'', ``img3'', ``norm2'', ``norm3'' and ``norm4'' were done. In total are 7 scans to determine $\sigma_{vis}$ \cite{pas_18}.\\
In the vdM scans the two beams were separated by $6 \sigma_{b} \approx 600 \mu$m and scanned across one another in a sequence of 25 steps with 30 seconds per step. For the beam imaging scans  one beam (beam 1 in the first pair, followed by beam 2 in the second pair) is kept fixed at its nominal position while the other is separated and scanned in 19 steps from  $+$4.5$\sigma_{b}$ to $-$4.5$\sigma_{b}$ with 46 seconds per step. In each of both scans (scan  pairs), the scan was performed first in the x direction and then in the y direction. \\For the background estimation, two Super Separation (SS) periods were carried out during the scan program, each 5 minutes long where the beams were separated by $6\sigma_{b}$ (where $\sigma_{b}$ the beam size) in both the x and y directions; the first SS period took place after norm3 and the second after norm4.\\
Fig. \ref{scan_prog} shows the beam positions for the two beams in the x and y directions as measured by the DOROS BPMs during the scan program, showing all scan pairs of the 2018 scan program.

\begin{center}
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=.45]{Chapter3/2018Scanprogram.png}
    \caption[2018 scan program]{Relative change in beam positions measured by the DOROS BPMs during the 2018 scan program for the two individual beams in the horizontal x and vertical y planes, as a function of the elapsed time from the beginning of the program. The top row shows the take 1 of the program before the alarm at CMS, while the bottom two rows show the program after the alarm \cite{pas_18}.}
    \label{scan_prog}
  \end{figure}
\end{center}

@@

The data recorded was processed with the CMS software (CMSSW) storing the data into a Hierarchical Data Format (HDF) with hd5 format, containing the times, beam positions, beam currents, and rates.\\
The data collected during the vdM scans is analysed with a software framework called ``vdM Framework'' (vdM FW), which is written in Python2 and uses the CERN’s data analysis package ``ROOT'' trough the ``PyRoot'' library. With the analysis and plotting tools, the vdM FW reads the hd5 file, producing some intermediate files: scan file, beam currents file, rates file and correction files, and applies some corrections to rates (Background, DynamicBeta), separation (OrbitDrift, LengthScale, BeamBeam) and beam currents (FBCT to DCCT calibration, Ghost and Satellite) and creates a graph file. This graph file contains the normalized rates (i.e. $R/N_{1}N_{2}$) and beam position. Each point correspond to the averaged rates in a time window of 30s and 46s for vdM scans (norm) and imaging scans, respectively. The points are then fitted with a predefined function to extract $\Sigma_{x,y}$ and peak values to compute $\sigma_{\text{vis}}$ for all the bunches in the scan.

%Description vdM FW:
%https://cds.cern.ch/record/2647770/files/CERN-THESIS-2018-245.pdf
%https://cds.cern.ch/record/2775639/files/CERN-THESIS-2021-086.pdf
%https://vdmdoc.docs.cern.ch/tutorial/tutorial.html
For some luminometers the rates are reprocessed, for example for backround subtracting. In the case of PCC the data was reprocessed for two purposes: i) the use of new data format and ii) a different selection of the pixel detector modules.\\
For PCC, the rates (and the corresponding times) were originally stored  per event into a ROOT file. Due the vdM FW procedures, the data analysis with the ROOT file carries some limitations, leading to the  use of only 1\% of the statistics. For this reason, the PCC data was reprocessed  to store the data into an hd5 files containing the rates stored per 1.32 s (NB4), allowing the use of the full statistics. This leads to a large improvement in the statistical uncertainty in the rates by a factor of 10, improved the fit model and the better determination of the beam parameters, relative to the previous analysis reported in ref. \cite{pas_18}.
%%\begin{center}
%%  \begin{figure}[ht]
%%    \centering
%%    \includegraphics[scale=.23]{Chapter4/fit_comparison.png}
%%    \caption[Comparison of statistics: example]{Single Gaussian fit of BCID 3380 from Y vdM scan (without corrections) using the old data format (left) and new data format (right). Figure used only to illustrate the difference and improvement of data formats described in the text.} 
%%    \label{fitcomp}
%%  \end{figure}
%%\end{center}
%%

\section{Module selection}
The implementation of a different selection of the modules in the pixel detector aims to remove instabilities in luminosity measurement. Previously, the same selection of modules was used for most of the year, for many run periods. Run periods corresponds to a division of time periods of the data set, for 2018 the calibration period is contained in the run period 2018-05-28 to 2018-07-07.\\
Using a different selection of modules per run period is more realistic as each period can have different number of good performing modules. The pixel detector conditions change over the time, for example by detector noise, aging effects and radiaiton damage.\\
The selection is done trough a module stability study. Module stability is based on variation of cluster count relative to total cluster count (module weight), root mean square (RMS) of weights distribution. The variations are considered in an interval of 23s (lumi section), which corresponds to the granularity (that is, the scale or level of detail in a set of data) of the luminosity database. Modules in barrel layer 1 are removed as these modules were not used in luminosity measurement. The module stability is re-evaluated (RMS values, 4\% RMS selection) by using iterative process until stable luminosity is attained. Fig. \ref{goodmodule} shows an example of a good and bad module.\\
For the calibration period, the module selection provided has 802 bad modules and 1054 good modules.\\
The results obtained from the analysis with the new selection of modules are shown below.
%: rms (root mean square) of weights distribution.  Module stability is re-evaluated by using iterative process until stable luminosity is attained \cite{ashish_stb1}.
% LS: it corresponds to the granularity (that is, the scale or level of detail in a set of data) of the luminosity database.
\begin{center}
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=.23]{Chapter4/good_module.png}
    \caption[Good-bad module stability profile example]{ Graphs showing stability profile for good and bad modules where weight is plotted as a function of lumi section.} %\cite{ashish_stb1}. } 
    \label{goodmodule}
  \end{figure}
\end{center}
%https://indico.unison.mx/event/152/
%https://indico.cern.ch/event/1125423/
%presentacion de sem inv ashish:
%https://indico.unison.mx/event/152/
%For the calibration period, the module selection provided has 802 bad modules and 1054 good modules.\\ %\cite{ashish_stb2}.\\%[https://indico.cern.ch/event/1157940/contributions/4862596/attachments/2437712/4176063/AshishSehrawat.pdf]
%The results obtained from the analysis with the new selection of modules are shown below.

\section{Background estimation}
To estimate the constant term in the measurements two super separation periods were carried out. Fig \ref{ssp_wide_bx1780} shows PCC per NB4 versus time for BCID 1780 to illustrate the two SS periods. The lower regions corresponds to the time window of 5 min where the beams were separate a distance of $6\sigma_{b}$, ensuring thatensuring that the only contribution is from background (either beam-induced background or electronics noise).
%\begin{center}
%\begin{figure}[ht]
%\centering
%\includegraphics[width=.49\textwidth]{Chapter4/plots/SS1_WIDE_BX1780.png}
%\includegraphics[width=.49\textwidth]{Chapter4/plots/SS2_WIDE_BX1780.png}
%\caption[Super Separation periods for BCID 1780]{PCC per NB4 versus time for BCID 1780. The lower regions corresponds to the super separation period I (left) and II (right).}
%\label{ssp_wide_bx1780}
%\end{figure}
%\end{center}
The background value in each BCID is estimated as the mean value of the distribution (or profile) of PCC per NB4. Fig \ref{ss1_hist_1780} shows the distribution of PCC per NB4 for BCID 1780 for SS period I, having a mean value of 0.322 with a standard error of the mean (SEM) of 0.018.

\newpage

\begin{center}
\begin{figure}[ht]
\centering
\includegraphics[width=.49\textwidth]{Chapter4/plots/SS1_WIDE_BX1780.png}
\includegraphics[width=.49\textwidth]{Chapter4/plots/SS2_WIDE_BX1780.png}
\caption[Super Separation periods for BCID 1780]{PCC per NB4 versus time for BCID 1780. The lower regions corresponds to the super separation period I (left) and II (right).}
\label{ssp_wide_bx1780}
\end{figure}
\end{center}

\begin{center}
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=.18]{Chapter4/plots/ss1_histo_bx1780.png}
    \caption[PCC per NB4 profile for BCID 1780 from SS period I]{ PCC per NB4 profile for BCID 1780 from SS period I.} 
    \label{ss1_hist_1780}
  \end{figure}
\end{center}

>>>>

@@

@@

\section{Module selection}

\section{Background estimation}

### 2023-02-16 tesis_Luis/Dedication/dedication.tex
Add files via upload

\begin{dedication}
\begin{center}
    \section*{Acknowledgements}
\end{center}
\justify{ 
Thanks to my family for their unconditional support and patience and for always being there when I needed them the most, without them I would not be here.\\
To my friends Carolina, Alejandro and Luisa, for all the help and for making my studies more fun.\\
To my advisor Dr. José Feliciano Benítez,  for always being there to answer my doubts, his patience, and for the opportunity of letting me work in the CMS collaboration. \\
Finally, thanks to CONACYT for supporting my studies with financial aid.}

>>>>

\begin{dedication}
\begin{center}
    \section*{Acknowledgements}
\end{center}
\justify{
